Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-02-08 constantRate/ConstCompFuzzExt.tex
revised claim d.3

We note that fuzzy extractors for the setting where $\Huse\le 0$ can be trivially constructed from a strong form of obfuscation, specifically, virtual black-box obfuscation for the class {\em proximity point programs}, $I_w(x)$ that tests if $x$ is within distance $t$ of $w$. However, currently we do not know if such obfuscation exists. Recent works constructed candidates indistinguishability obfuscators \cite{GargGH0SW13,PassTS13} and virtual black-box obfuscators in idealized model \cite{BrakerskiR13,BarakGKPS13} for all programs. However, these notions of obfuscation do not imply virtual-black-box obfuscation for proximity point programs in the plain model. The work of \cite{BarakBCKPS13} suggest an average-case virtual-black-box obfuscator for the family of {\em evasive} functions that test if a low-degree multi-variant arithmetic circuit evaluates to zero. However, we do not know if these functions include proximity point programs.

Both constructions are inspired by Canetti and Dakdouk's construction of digital lockers from point obfuscation~\cite{canetti2008obfuscating}.  Let $w=w_1 \dots w_\ell$, for $w_j\in \mathcal{Z}$. In the first construction (\consref{cons:first construction}), for each $j$, $\gen$ flips a coin $c_j$ and either obfuscates $I_{w_j}$ or picks a random point $r_j$ and obfuscates $I_{r_j}$.  This produces $\ell$ obfuscated programs $P_1,..., P_\ell$.  With a close value $w'$, $\rep$ then runs the obfuscated program $P_j$ on $w_j'$ and checks whether $P_j(w_j')=1$.  For most locations $j$, \rep can determine whether $w_j$ or a random value was obfuscated.  Thus, most bits of $c_j$ are recoverable. To tolerate errors,  the set of coins $c_1\dots c_\ell$  is chosen at random from the codewords of an error correcting code. This construction conducts entropy from $w$ to $c$.

@@

\begin{proof}
\noindent It suffices to show that for any two codewords that agree on $J^c$, the statistical distance is at most $(\ell-\beta)2^{-(\alpha+1)}$.
\begin{lemma}
\label{lem:codewords in I close}
Let $c^*$ be true value encoded in $X$ and let $c'$ a codeword in $C'$.  Then,
\[
\Delta( S^{I_X(\cdot, \cdot)}(c^*, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(c', 1^{\ell \log |Z|})) \le ( \ell -\beta) 2^{-(\alpha+1)}.
\]
\end{lemma}
\begin{proof}
Recall that for all $j\in J$, $\Hav(W_j | View(S))\geq \alpha$.  The only information about the correct value of $c_j^*$ is contained in the query responses.  When all responses are $0$ the view of $S$ is identical when presented with $c^*$ or $c'$.  We now show that for any value of $c^*$ all queries on $j \in J$ return $0$ with probability $1-2^{-\alpha+1}$.  Suppose not, that is suppose, the probability of at least one nonzero response on index $j$ is $> 2^{-(\alpha+1)}$.  Since $w, w'$ are independent of $r_j$, the probability of this happening when $c^*_j = 1$ is at most $q/2^n$ or equivalently $2^{-n+\log q}$.  Thus, it must occur with probability:
\begin{align}
2^{-\alpha+1}&<\Pr[\text{non zero response location }j]\nonumber \\
 &= \Pr[c_j^* =1]\Pr[\text{non zero response location }j\wedge c_j^*=1]\nonumber \\&+ \Pr[c_j^*=0] \Pr[\text{non zero response location }j \wedge c_j^*=0]\nonumber \\
&\le 1\times 2^{-n+\log q} + 1\times  \Pr[\text{non zero response location }j \wedge c_j^*=0] \label{eq:ways to remove ent}
\end{align}
We now show that $n-\log q \geq \alpha$:
\begin{claim}
\label{cl:ent bounded away from n}
$\alpha \le n-\log q$
\end{claim}
\begin{proof}
It suffices to show that there exists a simulator $S$ making $q$ queries such that the remaining entropy in any block is at most $n-\log q$.  Let $W_j$ be a distribution. Consider $S$ that asks the $q$ most likely outcomes. The total probability\lnote{what does total probability mean?} of this block\lnote{symbol?} must be at least $2^{-n+\log q}$.  After these queries the remaining min-entropy is at most:
\begin{align*}
\Hav(W_j | View(S)) &\leq  -\log \left(\Pr[\text{some }q_i=1]\times 1+ \Pr[\text{no }q_i=1]\times \Pr[\text{most likely outcome}|q_1,...,q_q]\right)\\
&\leq  -\log \left(\Pr[\text{some }q_i=1]\times 1\right)\\
&=-\log\left( 2^{-n+\log q} \right) = n-\log q
\end{align*}
This completes the proof of \clref{cl:ent bounded away from n}.
\end{proof}
\noindent
Rearranging terms in Equation~\ref{eq:ways to remove ent}, we have:
\begin{align*}
 \Pr[\text{non zero response location }j \wedge c_j=0] &>2^{-\alpha+1} - 2^{-(n-\log q)}=  2^{-\alpha}
 \end{align*}
 When there is a $1$ response and $c_j=0$ this means that there is no remaining min-entropy.  If this occurs with over $2^{-\alpha}$ probability this violates the block unguessability of $W$~(\defref{def:block guessable}).  By the union bound over the indices $j\in J$ the total probability of a $1$ in $J$ is at most $(\ell-\beta)2^{-\alpha+1}$. Recall that $c^*, c'$ match on all indices outside of $J$. Thus, for all $c^*, c'$ the statistical distance is at most $(\ell- \beta)2^{-\alpha+1}$.  This concludes the proof of \lemref{lem:codewords in I close}.
\end{proof}
By averaging over all points in $C'$ we conclude that $\Delta(S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(C', 1^{\ell \log |Z|})) < (\ell -\beta)2^{-(\alpha+1)}$.  This completes the proof of \lemref{lem:sim cannot distinguish}.
\end{proof}

>>>>

Both of our constructions are inspired by Canetti and Dakdouk's construction of digital lockers from point obfuscation~\cite{canetti2008obfuscating}.  Let $w=w_1 \dots w_\ell$, for $w_j\in \mathcal{Z}$. In the first construction (\consref{cons:first construction}), for each $j$, $\gen$ flips a coin $c_j$ and either obfuscates $I_{w_j}$ or picks a random point $r_j$ and obfuscates $I_{r_j}$.  This produces $\ell$ obfuscated programs $P_1,..., P_\ell$.  With a close value $w'$, $\rep$ then runs the obfuscated program $P_j$ on $w_j'$ and checks whether $P_j(w_j')=1$.  For most locations $j$, \rep can determine whether $w_j$ or a random value was obfuscated.  Thus, most bits of $c_j$ are recoverable. To tolerate errors,  the set of coins $c_1\dots c_\ell$  is chosen at random from the codewords of an error correcting code. This construction conducts entropy from $w$ to $c$.

@@

\begin{proof}
\noindent It suffices to show that for any two codewords that agree on $J^c$, the statistical distance is at most $(\ell-\beta)2^{-(\alpha+1)}$.
\begin{lemma}
\label{lem:codewords in I close}
Let $c^*$ be true value encoded in $X$ and let $c'$ a codeword in $C'$.  Then,
\[
\Delta( S^{I_X(\cdot, \cdot)}(c^*, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(c', 1^{\ell \log |Z|})) \le ( \ell -\beta) 2^{-(\alpha+1)}.
\]
\end{lemma}
\begin{proof}
Recall that for all $j\in J$, $\Hav(W_j | View(S))\geq \alpha$.  The only information about the correct value of $c_j^*$ is contained in the query responses.  When all responses are $0$ the view of $S$ is identical when presented with $c^*$ or $c'$.  We now show that for any value of $c^*$ all queries on $j \in J$ return $0$ with probability $1-2^{-\alpha+1}$.  Suppose not, that is suppose, the probability of at least one nonzero response on index $j$ is $> 2^{-(\alpha+1)}$.  Since $w, w'$ are independent of $r_j$, the probability of this happening when $c^*_j = 1$ is at most $q/\mathcal{Z}$ or  equivalently $2^{-\log |\mathcal{Z}|+\log q}$.  Thus, it must occur with probability:
\begin{align}
2^{-\alpha+1}&<\Pr[\text{non zero response location }j]\nonumber \\
 &= \Pr[c_j^* =1]\Pr[\text{non zero response location }j\wedge c_j^*=1]\nonumber \\&+ \Pr[c_j^*=0] \Pr[\text{non zero response location }j \wedge c_j^*=0]\nonumber \\
&\le 1\times 2^{-\log|\mathcal{Z}|+\log q} + 1\times  \Pr[\text{non zero response location }j \wedge c_j^*=0] \label{eq:ways to remove ent}
\end{align}
We now show that for an unguessable block source the remaining entropy $\alpha\leq \log |\mathcal{Z}|-\log q $:
\begin{claim}
\label{cl:ent bounded away from n}
If $W$ is a $(q, \alpha, \beta)$-block unguessable distribution over $\mathcal{Z}$ then $\alpha \le \log |\mathcal{Z}|-\log q$.
\end{claim}
\begin{proof}
\bnote{Leo I changed some of this under you.  It wasn't clear or probably right as written.}
\bnote{better name for guess, creativity is running low.}
Let $W$ be a $(q, \alpha, \beta)$-block unguessable distribution.  Let $J\subset\{1,..., \ell\}$ the set of good indices.
It suffices to show that there exists an $S$ making $q$ queries such that for some $j\in J, \Hav(W_j | S^{I_{W}(\cdot, \cdot)})\le \log |\mathcal{Z}| - \log q$.  Denote by $guess_{1}, ..., guess_q$ the $q$ most likely outcomes of $W_j$~(breaking ties arbitrarily).  Note that $\sum_{i=1}^q \Pr[W_j = guess_i]\geq q/|\mathcal{Z}|$.  Suppose not, this means that there is some $guess_i$ with probability $\Pr[W_j = guess_i] < 1/|\mathcal{Z}|$.  Since there are $\mathcal{Z} - q $ remaining possible values of $W_j$ for their total probability to be at least $1-q/|\mathcal{Z}|$ at least of these values has probability at least $1/\mathcal{Z}$.  This contradicts the statement $guess_1,..., guess_q$ are the most likely values.  Consider $S$ that queries its oracle on $(j, guess_1),.., (j, guess_q)$.  Denote by $Guess$ the random variable when $W_j\in \{guess_1,.., guess_q\}$  After these queries the remaining min-entropy is at most:
\begin{align*}
\Hav(W_j | S^{J_W(\cdot, \cdot)}) &=  -\log \left(\Pr[Guess=1]\times 1+ \Pr[Guess=0]\times \max_{w}\Pr[W_j = w|\not Guess =0]\right)\\
&\leq  -\log \left(\Pr[Guess=1]\times 1\right)\\
&=-\log\left( \frac{q}{|\mathcal{Z}|} \right) = \log|\mathcal{Z}|-\log q
\end{align*}
This completes the proof of \clref{cl:ent bounded away from n}.
\end{proof}
\noindent
Rearranging terms in Equation~\ref{eq:ways to remove ent}, we have:
\begin{align*}
 \Pr[\text{non zero response location }j \wedge c_j=0] &>2^{-\alpha+1} - 2^{-(\log |\mathcal{Z}|-\log q)}=  2^{-\alpha}
 \end{align*}
 When there is a $1$ response and $c_j=0$ this means that there is no remaining min-entropy.  If this occurs with over $2^{-\alpha}$ probability this violates the block unguessability of $W$~(\defref{def:block guessable}).  By the union bound over the indices $j\in J$ the total probability of a $1$ in $J$ is at most $(\ell-\beta)2^{-\alpha+1}$. Recall that $c^*, c'$ match on all indices outside of $J$. Thus, for all $c^*, c'$ the statistical distance is at most $(\ell- \beta)2^{-\alpha+1}$.  This concludes the proof of \lemref{lem:codewords in I close}.
\end{proof}
By averaging over all points in $C'$ we conclude that $\Delta(S^{I_X(\cdot, \cdot)}(C, 1^{\ell \log |Z|}), S^{I_X(\cdot, \cdot)}(C', 1^{\ell \log |Z|})) < (\ell -\beta)2^{-(\alpha+1)}$.  This completes the proof of \lemref{lem:sim cannot distinguish}.
\end{proof}
